---
title: "Formative Assessment 7"
author: "Ramilo, Zion John Yousef T."
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
  library(tidyverse)
```
a. In Example 16.3 with ðœ† = 4 per minute, use R to obtain:
    - P(T â‰¤ 0.25) = P(time between submissions is at most 15 seconds);
    - P(T > 0.5) = P(time between submissions is greater than 30 seconds);
    - P(0.25 < T < 1) = P(time between submissions is between 15 seconds and 1 minute).


P(T â‰¤ 0.25) = P(time between submissions is at most 15 seconds);
```{r}
cdf_exponential <- function(lambda,x){ #P(T <= t)
  probability <- 1-exp(-lambda*x)
  return (probability)
}

probability_submissions_15s <- round(cdf_exponential(4,0.25),2)
print(paste("Probability such that time between submissions is at most 15 seconds is",probability_submissions_15s))
```
P(T > 0.5) = P(time between submissions is greater than 30 seconds);
```{r}
#P(T > 0.5) = 1-P(T<=0.5)
probability_submissions_30s <- 1-round(cdf_exponential(4,0.5),2)
print(paste("Probability such that time between submissions is greater than 30 seconds",probability_submissions_30s))
```
P(0.25 < T < 1) = P(time between submissions is between 15 seconds and 1 minute).
```{r}
#P(0.25 < T < 1) = (P(T<=1) - P(T=1)) - P(T <= 0.25)
#Since P(T=1) = 0 we can say that P(T<1) = P(T<=1)
#So, P(0.25 < T < 1) = P(T<=1) - P(T <= 0.25)
probability_submissions_60s <- round(cdf_exponential(4,1),2)
print(paste("Probability such that time between submissions is at most 60 seconds",probability_submissions_60s))
print(paste("Probability such that the time between submissions is at least 15 seconds but at most 60 seconds",probability_submissions_60s-probability_submissions_15s))
```
b. The average rate of job submissions in a computer center is 2 per minute. If it can be assumed that the number of submissions per minute has a Poisson distribution, calculate the probability that:
    - more than two jobs will arrive in a minute;
    - at least 30 seconds will elapse between any two jobs;
    - less than 30 seconds will elapse between jobs;
    - a job will arrive in the next 30 seconds, if no jobs have arrived in the last 30 seconds.
    
More than two jobs will arrive in a minute;
```{r}
#P(X>2) =1- P(X <= 2) = 1-[P(X = 0)+P(X = 1)+P(X = 2)]

print(paste("The Probability that more than two jobs will arrive in a minute is equivalent to: ",round(1-sum(dpois(0:2,2)),2)))
```
At least 30 seconds will elapse between any two jobs;
```{r}
#P(X>=30) = P(X>=0.5) = 1-P(X<0.5) = 1-P(X<=0.5)-P(X=0.5) = 1-P(X<=0.5)
probability_elapsed_30s <- round(cdf_exponential(2,0.5),2)
print(paste("The Probability that at least 30 seconds will elapse between any two jobs: ",1-probability_elapsed_30s))
```
Less than 30 seconds will elapse between jobs;
```{r}
print(paste("The Probability that less than 30 seconds will elapse between jobs: ",probability_elapsed_30s))
```
A job will arrive in the next 30 seconds, if no jobs have arrived in the last 30
```{r}
#P(T<=60 | T>30) = P(T<=30)
probability_elapsed_30s <- round(cdf_exponential(2,0.5),2)
print(paste("The Probability that a job will arrive in the next 30 seconds, if no jobs have arrived in the last 30: ",probability_elapsed_30s))
```
c. A website receives an average of 15 visits per hour, which arrive following a
Poisson distribution.
    - Calculate the probability that at least 10 minutes will elapse without a visit.
    - What is the probability that in any hour, there will be less than eight visits?
    - Suppose that 15 minutes have elapsed since the last visit, what is the probability that a visit will occur in the next 15 minutes.
    - Calculate the top quartile, and explain what it means.

Calculate the probability that at least 10 minutes will elapse without a visit.
```{r}
#P(T >= 10) = 1-P(T<=10) = 1-P(T<10)-P(T=10) = 1-P(T<=10)
probability_elapsed_10m <- round(cdf_exponential(15,(1/6)),3)
print(paste("The Probability that the probability that at least 10 minutes will elapse without a visit.: ",1-probability_elapsed_10m))
```
What is the probability that in any hour, there will be less than eight visits?
```{r}
#P(X < 8) = P(X<=7) = [P(X = 0)+P(X = 1)+P(X = 2)+...+P(X=7)]
print(paste("The Probability that in any hour, there will be less than eight visits: ",round(sum(dpois(0:7,15)),3)))
```
Suppose that 15 minutes have elapsed since the last visit, what is the probability that a visit will occur in the next 15 minutes.
```{r}
#P(T<=30 | T>15) = P(T<=15)
probability_elapsed_15m <- round(cdf_exponential(15,(15/60)),2)
print(paste("The Probability a visit will occur in the next 15 minutes: ",probability_elapsed_15m))
```
Calculate the top quartile, and explain what it means.
```{r}
top_quartile <- qpois(0.75,15)
print(paste("Top quartile is equivalent to: ",top_quartile))
```
Top Quartile is the K such that it has a probability of 75% meaning it is the K that 75% of the time we are to see less than or equal to 18 visits those 25% are those that if it is larger than 18 visits it has a probability of 25%. 
