---
title: "Formative Assessment 8"
author: "Ramilo, Zion John Yousef T."
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
Question 1:
An analogue signal received at a detector, measured in microvolts, is normally
distributed with mean of 200 and variance of 256.

a. What is the probability that the signal will exceed 224 ğœ‡V?
```{r}
mean <- 200
variance <- 256
std_dev <- sqrt(variance)
x<- 224

q1_a <- 1-pnorm(x,mean = mean,sd=std_dev)
cat("P(X  > 224) =",q1_a,"or",round(q1_a,4)*100,"%")
```
b. What is the probability that it will be between 186 and 224 ğœ‡V?
```{r}

x_1 <- 224
x_2 <- 186

q1_b_1 <- pnorm(x_1,mean = mean,sd=std_dev)
q1_b_2 <- pnorm(x_2,mean = mean,sd=std_dev)

q1_b <- q1_b_1-q1_b_2
cat("P(186 < X < 224) =",q1_b,"or",round(q1_b,4)*100,"%")
```
c. What is the micro voltage below which 25% of the signals will be?
```{r}
q1_c <- qnorm(0.25,mean = mean,sd=std_dev)
cat("P(X < x) =",0.25,"equates to:",q1_c,"micro volts")
```
d. What is the probability that the signal will be less than 240 ğœ‡V, given that
it is larger than 210 ğœ‡V?
```{r}
x_1 <- 240
x_2 <- 210

q1_d_1 <- pnorm(x_1,mean = mean,sd=std_dev)
q1_d_2 <- pnorm(x_2,mean = mean,sd=std_dev)

q1_d_numerator <- q1_d_1-q1_d_2
q1_d_dinominator <- 1-q1_d_2

q1_d <- q1_d_numerator/q1_d_dinominator

cat("P(210 < X < 240)/ P(X > 210) =",q1_d,"or",round(q1_d,4)*100,"%")
```
e. Estimate the interquartile range.
```{r}
quartile_3 <- 0.75
quartile_1 <- 0.25

q1_e_1 <- qnorm(quartile_3,mean = mean,sd=std_dev)
q1_e_2 <- qnorm(quartile_1,mean = mean,sd=std_dev)

cat("Interquartile Range: ",q1_e_1-q1_e_2,"micro volts")
```
f. What is the probability that the signal will be less than 220 ğœ‡V, given that
it is larger than 210 ğœ‡V?
```{r}
x_1 <- 220
x_2 <- 210

q1_f_1 <- pnorm(x_1,mean = mean,sd=std_dev)
q1_f_2 <- pnorm(x_2,mean = mean,sd=std_dev)

q1_f_numerator <- q1_f_1-q1_f_2
q1_f_dinominator <- 1-q1_f_2

q1_f <- q1_f_numerator/q1_f_dinominator

cat("P(210 < X < 220)/ P(X > 210) =",q1_f,"or",round(q1_f,4)*100,"%")
```
g.  If we know that a received signal is greater that 200 ğœ‡V, what is the probability that is in fact greater than 220 ğœ‡V?
```{r}
x_1 <- 220
x_2 <- 200

q1_g_1 <- 1-pnorm(x_1,mean = mean,sd=std_dev)
q1_g_2 <- 1-pnorm(x_2,mean = mean,sd=std_dev)

q1_g <- q1_g_1/q1_g_2

cat("P(X > 220)/ P(X > 200) =",q1_g,"or",round(q1_g,4)*100,"%")
```
Question 2:
A manufacturer of a particular type of computer system is interested in improving its customer support services. As a first step, its marketing department has
been charged with the responsibility of summarizing the extent of customer
problems in terms of system failures. Over a period of six months, customers
were surveyed and the amount of downtime (in minutes) due to system failures
they had experienced during the previous month was collected. The average
downtime was found to be 25 minutes and a variance of 144. If it can be
assumed that downtime is normally distributed:

(a) obtain bounds which will include 95% of the downtime of all the customers;
```{r}
confidence_lvl <- 0.95
two_tailed_test <- (1-confidence_lvl)/2

upperBound <- qnorm(confidence_lvl+two_tailed_test,mean = 25,sd = sqrt(144))
lowerBound <- qnorm(two_tailed_test,mean = 25,sd = sqrt(144))

cat("Bounds includes: ","(",lowerBound,"minutes ,",upperBound,"minutes )")
```
(b) obtain the bound above which 10% of the downtime is included.
```{r}
minimumBound <- qnorm(0.9,mean = 25,sd = sqrt(144))
cat("Top 10% of the downtime is included: ",minimumBound,"minutes")
```




