---
title: "Formative Assessment 6"
author: "Zion John Yousef T. Ramilo"
date: "2024-02-25"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

  I. Geometric Distribution. Provide an R code for the geometric distribution. The geometric distribution is a probability distribution that models the number of trials required to achieve the first success in a sequence of Bernoulli trials, where each trial has a constant probability of success.
  
1. Set the probability of success:  p <- 0.2
2. Generate 1000 random variables from the geometric distribution.
```{r}
geom_dist <- rgeom(1000,0.2)
geom_dist_df <- data.frame("n_failures_before_success" = geom_dist)
head(geom_dist_df)
tail(geom_dist_df)
```
3. Calculate some basic statistics:
  - mean_x <- mean(x)
  - var_x <- var(x)
  - sd_x <- sd(x)
4. Print the results in item 3 with the following output (string): 
  - Number of trials required to achieve first success:
  - Mean (in 2 decimal places):
  - Variance (in 2 decimal places):
  - Sandard deviation ( in 2 decimal places):
```{r}
geom_dist_df_summary <- geom_dist_df %>% 
  summarize("Mean" = round(mean(n_failures_before_success),2),
            "Variance" = round(sd(n_failures_before_success),2),
            "StandardDiviation" = round(var(n_failures_before_success),2))
first_row <- paste("Number of trials required to achieve first success: ",1000)
second_row <- paste("Mean (in 2 decimal places): ",geom_dist_df_summary$Mean)
third_row <- paste("Variance (in 2 decimal places):",geom_dist_df_summary$Variance)
fourth_row <- paste("Standard deviation ( in 2 decimal places):",geom_dist_df_summary$StandardDiviation)
cat("\n",first_row, "\n", second_row, "\n", third_row, "\n", fourth_row, "\n")
```
5. Plot the histogram of the results.
```{r}
ggplot(data = geom_dist_df,mapping = aes(x=n_failures_before_success))+
  geom_histogram(binwidth = 1)+
    labs(x="Number of Failures Before Successes",
         y="Frequency",
         title = "Histogram of a Geometric Distribution",
         subtitle = "1000 Trials")
```
II. Hypergeometric Distribution.  Consider a plant manufacturing IC chips of which 10% are expected to be defective. The chips are packed in boxes for export. Before transportation, a sample is drawn from each box. Estimate the probability that the sample contains more than 10% defectives, when:
```{r}
hypergeom_prob <- function(success,n_success,n,N){
  M <- success
  L <- N-M
  x <- n_success
  x_ <- n-x
  numerator <- choose(M,x) * choose(L,x_)
  dinominator <- choose(N,n)
  probability <- numerator / dinominator
  return(probability)
}
```


  1. A sample of 10 is selected from a box of 40;
```{r}
estimated_success_40 <- 40*0.1
popu_40 <- 1-hypergeom_prob(estimated_success_40,1,10,40)
```
  2. A sample of 10 is selected from a box of 5000.
```{r}
estimated_success_5000 <- 5000*0.1
popu_5000 <- 1-hypergeom_prob(estimated_success_5000,1,10,5000)

cat("\n","Estimated the probability that more than 10% of the sample is defective from a population of 40: ",popu_40, "\n","Estimated the probability that more than 10% of the sample is defective from a population of 5000: ", popu_5000)
```

